\documentclass[../draft_1.tex]{subfiles}

\begin{document}


\chapter{Conclusions and Outlook}

In this thesis we developed a least-squares space--time finite element discretisation for parabolic reaction-diffusion systems with potentially nonlinear forcing terms. We implemented a version of this solution method that can handle test cases of one space dimension. It uses a uniform mesh with rectangular elements for the discretisation of the space--time domain and the finite--dimensional approximation space $Q_1$, that is spanned by bilinear polynomials which are evaluated at the nodes of the mesh cells. To solve the arising linear systems of equations, we implemented a linear geometric multigrid V--cycle algorithm, that uses \textit{equal} space--time coarsening and two different Vanka--type smoothers, a block--Jacobi relaxation scheme and a 2--colouring Gauss--Seidel line smoother. For solving the nonlinear problem we implemented a trust--region dogleg method. 
\smallskip
\\
We did not really know how the solution method would perform on our test cases as to our knowledge there is no literature to compare our results to for space--time least-squares finite element discretisations. This applies for both, the multigrid algorithm as well as the energy minimisation through the trust--region method. 
\smallskip 
\\
As we tested the multigrid algorithm on sparse, symmetric, positive definite matrices arising from the discretisation of a mixed system of equations of rather small size we had hoped for better convergence rates and more consistent scaling. However this was not really the case. The multigrid algorithm converged for all test cases but at unsatisfactory convergence rates and it also did not scale to an satisfactory extent. After a detailed error analysis of various parameters and an eigenvalue decomposition of the system we can conclude the following. 
\smallskip
\\
We then tried to understand better what affected the rate of convergence.  


\begin{itemize}
	\item really not a straight forward problem, dependent on the many parameters ... 
	\item maybe the scaling ... how terms are weighted ... or this multigrid thing, since the algorithm seems to work alright for other problems ... unfortunate combination of things
	\item so many more things what to test, summarise what worked best. what seems to not work well at all
	\item really absolutely toy examples, but important understand behavior there before working with really large systems
	
\end{itemize}

It is more complicated to understand the influence of the diffusion constant $d$ the term $\nabla u$ is coupled with $\sigma$ and therefore influences the overall dynamics. The scaling parameter $c_2$ however affects only the second term. And we observed a significant speed up in the convergence rate by increasing this term. Probably due to the fact that the increase in $c_2$ leads to an increase of the small eigenvalues of $H$ which in turn then lowers the condition numer $\kappa(H)$ and therefore leads to faster convergence. 


We additionally saw that the multigrid algorithm seemed to perform well for the one-dimensional Laplace equation. Therefore it seems a reasonable hypothesis that the time derivative in $u$ is causing problems. It is known that multigrid methods do not achieve the same good performance results for parabolic problems as they do for elliptic ones [source]. And many strategies are being developed to modify or construct the arising systems of equations to enhance the performance of the multigrid algorithm, e.g. in \cite{langer2017multipatch}. It is often necessary to introduce some sort of stabilisation. The potential problem we might be experiencing here might have to do with the fact that we do not have a time derivative in $\sigma$. All first order derivatives of $u$ are present in the equation, however as $\sigma$ is defined as the gradient of $u$ in space it is generally still a function of $t$, without containing 

Therefore another possibility could be to choose different basis functions for $\sigma$ that are constant in time, in order to avoid the potential instability arising from this. 


Additionally, further investigation into theoretical framework. Find proof or counterexample of coercivity such that we get existence and uniqueness as well as convergence of the discretisation to actual solution. Or need to choose spaces differently. 
\end{document}